---
title: "Bloom: 3-Level Mixed Effects Model"
subtitle: "Crazy Poor Bayesians"
author: "Omar Shaikh, Shardul Sapkota"
date: "`r Sys.time()`"
urlcolor: blue
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
```

# Overview

This document fits a 3-level linear mixed-effects model (days nested within weeks nested within participants) to each of the four HealthKit wearable outcomes: step count, active energy burned, exercise time, and distance walking/running.

The model structure is:

$$Y \sim S + S\text{:condition} + S\text{:week} + S\text{:condition:week} + \text{dow} + (1|\text{uid/week})$$

We test:

- **H1**: Overall baseline-to-study increase ($S$)
- **H2**: Condition difference in the baseline-to-study shift ($S$:condition)
- **H3**: Condition difference in the weekly slope ($S$:condition:week)

# Setup

```{r setup-libs, message=FALSE, warning=FALSE}
library(nlme)
library(lme4)
library(lmerTest)
library(tidyverse)
library(glue)
library(effectsize)
```

# Data

```{r load-data}
df <- read_csv("data/daily_hk.csv", col_types = cols()) %>%
  mutate(
    uid       = factor(uid),
    condition = if_else(group == "treatment", 1L, 0L),
    S         = if_else(day > 0, 1L, 0L),
    d_study   = if_else(day > 0, day, 0L),

    week = case_when(
      day <= 0 ~ 0L,
      TRUE     ~ ((day - 1L) %/% 7L) + 1L
    ) %>% factor(levels = 0:4),
    week_num = as.integer(as.character(week)),

    dow_idx = ((day - 1L) %% 7L + 7L) %% 7L + 1L,
    dow = factor(
      dow_idx,
      levels = 1:7,
      labels = c("Sunday","Monday","Tuesday","Wednesday",
                 "Thursday","Friday","Saturday")
    )
  ) %>%
  filter(day <= 28)
```

# Model Fitting

We loop over each outcome, apply inclusion criteria (≥30 non-missing baseline days; <3 missing study days), fit the 3-level model, and collect coefficient estimates for the three hypotheses.

```{r models, warning=FALSE}
vars <- c(
  "hk_stepcount",
  "hk_activeenergyburned",
  "hk_appleexercisetime",
  "hk_distancewalkingrunning"
)

res <- list()

for (v in vars) {
  dat <- df %>%
    select(uid, condition, S, d_study, week, week_num, dow, day, y = all_of(v))

  ## baseline inclusion: >= 30 baseline days non-missing
  keep_uid_baseline <- dat %>%
    filter(day <= 0) %>%
    group_by(uid) %>%
    summarise(n_base = sum(!is.na(y)), .groups = "drop") %>%
    filter(n_base >= 30) %>%
    pull(uid)

  dat <- dat %>% filter(uid %in% keep_uid_baseline)

  ## study inclusion: exclude participants with >= 3 missing study days
  keep_uid_study <- dat %>%
    filter(day >= 1, day <= 28) %>%
    group_by(uid) %>%
    summarise(n_missing_study = sum(is.na(y)), .groups = "drop") %>%
    filter(n_missing_study < 3) %>%
    pull(uid)

  dat <- dat %>% filter(uid %in% keep_uid_study)

  fit <- lme(
    y ~ S + S:condition + S:week_num + S:condition:week_num + dow,
    random   = ~ 1 | uid/week,
    weights  = varIdent(form = ~ 1 | uid),
    data     = dat,
    method   = "REML",
    na.action = na.omit,
    control  = lmeControl(maxIter = 100,
                          msMaxIter = 300,
                          msMaxEval = 5000)
  )

  cat("\n--- 3-level model:", v, "---\n")
  print(summary(fit))

  coefs  <- summary(fit)$tTable
  terms  <- c("S", "S:condition", "S:condition:week_num")
  hypo   <- c("H1","H2","H3")
  for (i in seq_along(terms)) {
    term  <- terms[i]
    t_val <- coefs[term, "Value"] / coefs[term, "Std.Error"]
    df_er <- coefs[term, "DF"]
    d_val <- t_to_d(t_val, df_error = df_er, hedges.correction = FALSE)$d

    res[[length(res) + 1]] <- tibble(
      outcome    = v,
      hypothesis = hypo[i],
      term       = term,
      estimate   = coefs[term, "Value"],
      se         = coefs[term, "Std.Error"],
      df         = coefs[term, "DF"],
      cohen_d    = d_val,
      p_raw      = coefs[term, "p-value"]
    )
  }
}
```

# Results

```{r results}
results_df <- bind_rows(res) %>%
  mutate(p_holm = p.adjust(p_raw, method = "holm"))

print(results_df)
```

Holm-corrected p-values (`p_holm`) are reported to control the familywise error rate across the four outcomes.
