---
title: "Bloom: Exercise Time Adherence (150 min/week)"
subtitle: "Crazy Poor Bayesians"
author: "Omar Shaikh, Shardul Sapkota"
date: "`r Sys.time()`"
urlcolor: blue
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
```

# Overview

This document examines adherence to the WHO-recommended 150 minutes of moderate physical activity per week using the `hk_appleexercisetime` variable. We report the percentage of participants meeting this threshold at baseline vs. during the study, overall and broken down by condition.

# Setup

```{r setup-libs, message=FALSE, warning=FALSE}
library(tidyverse)
library(glue)
```

# Data

```{r load-data}
df <- read_csv("data/daily_hk.csv", col_types = cols()) %>%
  mutate(
    uid       = factor(uid),
    condition = if_else(group == "treatment", 1L, 0L),
    S         = if_else(day > 0, 1L, 0L),
    d_study   = if_else(day > 0, day, 0L),

    week = case_when(
      day <= 0 ~ 0L,
      TRUE     ~ ((day - 1L) %/% 7L) + 1L
    ) %>% factor(levels = 0:4),
    week_num = as.integer(as.character(week)),

    dow_idx = ((day - 1L) %% 7L + 7L) %% 7L + 1L,
    dow = factor(
      dow_idx,
      levels = 1:7,
      labels = c("Sunday","Monday","Tuesday","Wednesday",
                 "Thursday","Friday","Saturday")
    )
  ) %>%
  filter(day <= 28)
```

# Inclusion Criteria

We apply the same inclusion criteria as the main analysis: ≥30 non-missing baseline days and <3 missing study days.

```{r inclusion}
exercise_time_data <- df %>%
  select(uid, group, condition, day, week, y = hk_appleexercisetime)

keep_uid_baseline_ex <- exercise_time_data %>%
  filter(day <= 0) %>%
  group_by(uid) %>%
  summarise(n_base = sum(!is.na(y)), .groups = "drop") %>%
  filter(n_base >= 30) %>%
  pull(uid)

exercise_time_data <- exercise_time_data %>% filter(uid %in% keep_uid_baseline_ex)

keep_uid_study_ex <- exercise_time_data %>%
  filter(day >= 1, day <= 28) %>%
  group_by(uid) %>%
  summarise(n_missing_study = sum(is.na(y)), .groups = "drop") %>%
  filter(n_missing_study < 3) %>%
  pull(uid)

exercise_time_data <- exercise_time_data %>% filter(uid %in% keep_uid_study_ex)
```

# Weekly Minutes and Adherence

We average daily exercise time within each participant-week, scale to weekly minutes, then classify participants as meeting the 150 min/week threshold within each period (baseline vs. study).

```{r adherence}
exercise_time_data <- exercise_time_data %>%
  mutate(
    week_num = as.integer(as.character(week)),
    period   = if_else(week_num <= 0, "baseline", "study")
  )

weekly_minutes <- exercise_time_data %>%
  group_by(uid, condition, group, period, week_num) %>%
  summarise(
    weekly_minutes = 7 * mean(y, na.rm = TRUE),  # corrects for weeks with < 7 days
    .groups = "drop"
  )

person_period <- weekly_minutes %>%
  group_by(uid, condition, group, period) %>%
  summarise(
    avg_weekly_minutes = mean(weekly_minutes, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(meets_150 = avg_weekly_minutes >= 150)
```

# Results

## Overall (baseline vs. study)

```{r overall}
overall <- person_period %>%
  group_by(period) %>%
  summarise(
    n_participants = n(),
    n_meeting      = sum(meets_150, na.rm = TRUE),
    pct_meeting    = round(100 * n_meeting / n_participants, 1),
    .groups = "drop"
  ) %>%
  arrange(match(period, c("baseline", "study")))

print(overall)
```

## By Condition

```{r by-condition}
by_condition <- person_period %>%
  group_by(period, group) %>%
  summarise(
    n_participants = n(),
    n_meeting      = sum(meets_150, na.rm = TRUE),
    pct_meeting    = round(100 * n_meeting / n_participants, 1),
    .groups = "drop"
  ) %>%
  arrange(match(period, c("baseline", "study")), group)

print(by_condition)
```
