---
title: "Bloom: Visualizations"
subtitle: "Crazy Poor Bayesians"
author: "Omar Shaikh, Shardul Sapkota"
date: "`r Sys.time()`"
urlcolor: blue
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath("../.."))
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4)
```

```{r libs}
library(tidyverse)
library(glue)
```

```{r theme}
theme_set(theme_classic(base_size = 12))
condition_colors  <- c("Control" = "#4C9BE8", "Treatment" = "#E87B4C")
condition_fills   <- c("Control" = "#4C9BE8", "Treatment" = "#E87B4C")
```

# Data

```{r load-data}
df_hk <- read_csv("data/daily_hk.csv", col_types = cols()) %>%
  mutate(
    uid       = factor(uid),
    Condition = if_else(group == "treatment", "Treatment", "Control"),
    condition = if_else(group == "treatment", 1L, 0L),
    S         = if_else(day > 0, 1L, 0L),
    week_num  = case_when(
      day <= 0 ~ 0L,
      TRUE     ~ as.integer(((day - 1L) %/% 7L) + 1L)
    )
  ) %>%
  filter(day <= 28)

df_pre_post <- read_csv("data/pre_post.csv", col_types = cols()) %>%
  mutate(
    Condition = if_else(condition == "T", "Treatment", "Control"),
    uid       = row_number()   # no UID column; use row index
  )

df_app <- read_csv("data/daily_hk_app_usage.csv", col_types = cols()) %>%
  mutate(
    uid       = factor(uid),
    Condition = if_else(condition == "treatment", "Treatment", "Control")
  ) %>%
  filter(day >= 1, day <= 28)
```

```{r inclusion}
# Apply same inclusion criteria as main models (≥30 baseline days; <3 missing study days)
keep_uid <- df_hk %>%
  group_by(uid) %>%
  summarise(
    n_base         = sum(!is.na(hk_stepcount) & day <= 0),
    n_missing_study = sum(is.na(hk_stepcount) & day >= 1 & day <= 28),
    .groups = "drop"
  ) %>%
  filter(n_base >= 30, n_missing_study < 3) %>%
  pull(uid)

df_hk_inc <- df_hk %>% filter(uid %in% keep_uid)
```

---

# Plot 1 — Physical Activity Over Time by Condition

Mean daily step count aggregated by week (week 0 = baseline, weeks 1–4 = study period).
Error bars show ±1 SE across participants.

```{r plot1-steps-over-time}
week_means <- df_hk_inc %>%
  group_by(uid, Condition, week_num) %>%
  summarise(mean_steps = mean(hk_stepcount, na.rm = TRUE), .groups = "drop") %>%
  group_by(Condition, week_num) %>%
  summarise(
    m  = mean(mean_steps, na.rm = TRUE),
    se = sd(mean_steps, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  mutate(week_label = if_else(week_num == 0, "Baseline", glue("Week {week_num}")),
         week_label = factor(week_label, levels = c("Baseline", glue("Week {1:4}"))))

ggplot(week_means, aes(x = week_label, y = m, color = Condition, group = Condition)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0.2, linewidth = 0.7) +
  geom_vline(xintercept = 1.5, linetype = "dashed", color = "gray50", linewidth = 0.6) +
  scale_color_manual(values = condition_colors) +
  labs(
    title   = "Mean Daily Step Count by Week and Condition",
    x       = NULL,
    y       = "Mean daily steps",
    color   = NULL
  ) +
  theme(legend.position = "top")
```

```{r plot1-exercise-time}
week_means_ex <- df_hk_inc %>%
  group_by(uid, Condition, week_num) %>%
  summarise(mean_ex = mean(hk_appleexercisetime, na.rm = TRUE), .groups = "drop") %>%
  group_by(Condition, week_num) %>%
  summarise(
    m  = mean(mean_ex, na.rm = TRUE),
    se = sd(mean_ex, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  mutate(week_label = if_else(week_num == 0, "Baseline", glue("Week {week_num}")),
         week_label = factor(week_label, levels = c("Baseline", glue("Week {1:4}"))))

ggplot(week_means_ex, aes(x = week_label, y = m, color = Condition, group = Condition)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0.2, linewidth = 0.7) +
  geom_vline(xintercept = 1.5, linetype = "dashed", color = "gray50", linewidth = 0.6) +
  scale_color_manual(values = condition_colors) +
  labs(
    title   = "Mean Daily Exercise Time (min) by Week and Condition",
    x       = NULL,
    y       = "Mean daily exercise time (min)",
    color   = NULL
  ) +
  theme(legend.position = "top")
```

---

# Plot 2 — Baseline-to-Study Change by Condition

Participant-level change scores (study mean − baseline mean) for all four HealthKit outcomes.

```{r plot2-change-scores}
outcome_labels <- c(
  hk_stepcount              = "Step Count",
  hk_activeenergyburned     = "Active Energy (kcal)",
  hk_appleexercisetime      = "Exercise Time (min)",
  hk_distancewalkingrunning = "Distance (km)"
)

change_scores <- df_hk_inc %>%
  pivot_longer(
    cols      = c(hk_stepcount, hk_activeenergyburned,
                  hk_appleexercisetime, hk_distancewalkingrunning),
    names_to  = "outcome",
    values_to = "value"
  ) %>%
  group_by(uid, Condition, outcome, S) %>%
  summarise(period_mean = mean(value, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = S, values_from = period_mean, names_prefix = "S") %>%
  mutate(change = S1 - S0,
         outcome = factor(outcome, levels = names(outcome_labels),
                          labels = outcome_labels))

ggplot(change_scores, aes(x = Condition, y = change, fill = Condition)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray60") +
  geom_boxplot(alpha = 0.7, outlier.shape = 21, outlier.size = 1.5,
               width = 0.5) +
  scale_fill_manual(values = condition_fills) +
  facet_wrap(~ outcome, scales = "free_y", nrow = 1) +
  labs(
    title = "Baseline-to-Study Change Scores by Condition",
    x     = NULL,
    y     = "Change (study − baseline daily mean)",
    fill  = NULL
  ) +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text = element_text(size = 9))
```

---

# Plot 3 — Pre–Post Psychological Change by Condition

Dumbbell plot showing mean pre and post scores by condition for key psychological outcomes.
Normalized composite scores are rescaled to a 1–5 Likert-style axis (`norm × 4 + 1`).

```{r plot3-pre-post}
psych_vars <- tribble(
  ~pre_col,                                    ~post_col,                                    ~label,                        ~scale,
  "adequacy_mindset_total_norm_pre",           "adequacy_mindset_total_norm_post",           "PA Adequacy\nMindset",        "norm01",
  "process_mindset_exercise_total_norm_pre",   "process_mindset_exercise_total_norm_post",   "PA Process\nMindset",         "norm01",
  "satisfaction_physical_activity_pre",        "satisfaction_physical_activity_post",        "Satisfaction:\nPA Level",     "likert7",
  "satisfaction_physical_health_pre",          "satisfaction_physical_health_post",          "Satisfaction:\nPhysical Health", "likert7"
)

# norm01: 0-1 normalized → rescale to 1-5
# likert7: raw 1-7 item → rescale to 1-5
rescale_col <- function(vals, scale) {
  if (scale == "norm01")   vals * 4 + 1
  else if (scale == "likert7") (vals - 1) / 6 * 4 + 1
  else vals
}

psych_means <- psych_vars %>%
  pmap_dfr(function(pre_col, post_col, label, scale) {
    df_pre_post %>%
      group_by(Condition) %>%
      summarise(
        pre_m  = mean(rescale_col(.data[[pre_col]],  scale), na.rm = TRUE),
        post_m = mean(rescale_col(.data[[post_col]], scale), na.rm = TRUE),
        .groups = "drop"
      ) %>%
      mutate(label = label)
  }) %>%
  mutate(
    label     = factor(label, levels = rev(psych_vars$label)),
    delta     = post_m - pre_m,
    delta_lbl = sprintf("%+.2f", delta),
    # vertical offset so Treatment and Control don't overlap
    y_offset  = if_else(Condition == "Treatment", 0.18, -0.18),
    y_pos     = as.numeric(label) + y_offset
  )

cond_colors <- c("Treatment" = "#2a7a2a", "Control" = "#aaaaaa")

ggplot(psych_means) +
  # arrow from pre → post
  geom_segment(
    aes(x = pre_m, xend = post_m, y = y_pos, yend = y_pos,
        color = Condition),
    arrow = arrow(length = unit(0.18, "cm"), type = "closed"),
    linewidth = 0.9
  ) +
  # dot at pre
  geom_point(aes(x = pre_m, y = y_pos, color = Condition), size = 3) +
  # label: "Pre → Post (Δ)"
  geom_text(
    aes(x = post_m, y = y_pos,
        label = sprintf("Pre: %.2f → Post: %.2f (Δ=%s)", pre_m, post_m, delta_lbl),
        color = Condition,
        hjust = if_else(post_m >= pre_m, -0.08, 1.08)),
    size = 2.8, show.legend = FALSE
  ) +
  scale_color_manual(values = cond_colors,
                     guide  = guide_legend(reverse = TRUE)) +
  scale_y_continuous(
    breaks = seq_along(levels(psych_means$label)),
    labels = levels(psych_means$label)
  ) +
  scale_x_continuous(limits = c(1, 6.5),
                     breaks = 1:5,
                     labels = c("1\nStrongly\nDisagree", "2", "3", "4",
                                "5\nStrongly\nAgree")) +
  labs(
    title  = "Mindset & Satisfaction: Pre–Post by Condition",
    x      = "Average Score (Likert Scale)",
    y      = NULL,
    color  = NULL
  ) +
  theme_classic(base_size = 11) +
  theme(
    legend.position   = "top",
    axis.text.y       = element_text(hjust = 1, lineheight = 0.9),
    panel.grid.major.x = element_line(color = "gray88", linetype = "dashed"),
    plot.title        = element_text(face = "bold", hjust = 0.5)
  )
```

---

# Plot 4 — App Engagement (Treatment Group)

Distribution of daily chat interactions and app usage during the study period,
shown for treatment participants only (control had no AI coaching features).

```{r plot4-engagement}
app_treat <- df_app %>%
  filter(Condition == "Treatment") %>%
  group_by(uid) %>%
  summarise(
    total_chats    = sum(chat_counts, na.rm = TRUE),
    total_messages = sum(chat_user_messages_count, na.rm = TRUE),
    total_duration = sum(app_duration, na.rm = TRUE) / 60,  # convert to minutes
    .groups        = "drop"
  )

p_chats <- ggplot(app_treat, aes(y = total_chats)) +
  geom_boxplot(fill = condition_fills["Treatment"], alpha = 0.7,
               width = 0.4, outlier.shape = 21) +
  labs(title = "Total Chat Sessions", y = "# sessions", x = NULL) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

p_msgs <- ggplot(app_treat, aes(y = total_messages)) +
  geom_boxplot(fill = condition_fills["Treatment"], alpha = 0.7,
               width = 0.4, outlier.shape = 21) +
  labs(title = "Total User Messages", y = "# messages", x = NULL) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

p_dur <- ggplot(app_treat, aes(y = total_duration)) +
  geom_boxplot(fill = condition_fills["Treatment"], alpha = 0.7,
               width = 0.4, outlier.shape = 21) +
  labs(title = "Total App Duration", y = "minutes", x = NULL) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

# Arrange side by side using patchwork if available, otherwise cowplot
if (requireNamespace("patchwork", quietly = TRUE)) {
  library(patchwork)
  p_chats + p_msgs + p_dur +
    plot_annotation(title = "App Engagement (Treatment Group, Study Period)")
} else {
  print(p_chats)
  print(p_msgs)
  print(p_dur)
}
```

```{r plot4-daily-chat-trend}
# Daily mean chat sessions by week for treatment group
daily_chat <- df_app %>%
  filter(Condition == "Treatment") %>%
  mutate(
    week_num  = as.integer(ceiling(day / 7)),
    week_num  = pmin(week_num, 4),
    week_label = glue("Week {week_num}") %>%
      factor(levels = glue("Week {1:4}"))
  ) %>%
  group_by(uid, week_label) %>%
  summarise(weekly_chats = sum(chat_counts, na.rm = TRUE), .groups = "drop") %>%
  group_by(week_label) %>%
  summarise(
    m  = mean(weekly_chats, na.rm = TRUE),
    se = sd(weekly_chats, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(daily_chat, aes(x = week_label, y = m, group = 1)) +
  geom_line(color = condition_colors["Treatment"], linewidth = 0.9) +
  geom_point(color = condition_colors["Treatment"], size = 2.5) +
  geom_errorbar(aes(ymin = m - se, ymax = m + se),
                color = condition_colors["Treatment"], width = 0.2, linewidth = 0.7) +
  labs(
    title = "Weekly Chat Sessions Over Time (Treatment Group)",
    x     = NULL,
    y     = "Mean chat sessions per week"
  )
```
