---
title: "Bloom: Power Analysis"
subtitle: "Crazy Poor Bayesians"
author: "Omar Shaikh, Shardul Sapkota"
date: "`r Sys.time()`"
urlcolor: blue
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
```

# Overview

This document reports the power analysis for the Bloom study. We compute the sample sizes required to achieve 80% power under small, medium, and large effect sizes, and estimate the minimally detectable effect sizes (MDES) given the actual sample.

# Required Sample Sizes for 80% Power

For **H1** (baseline vs. study, within-person paired comparison) and **H2** (LLM vs. control, between-group comparison), we compute the required N at $d \in \{0.3, 0.5, 0.8\}$, $\alpha = .05$.

```{r sample-sizes}
library(pwr)
library(tidyverse)

## Sensitivity Analysis
ds <- c(0.3, 0.5, 0.8)

## ---- H1: study > baseline (paired) ----
h1_required <- sapply(ds, function(d) {
  out <- pwr.t.test(
    d = d,
    sig.level = 0.05,
    power = 0.80,
    type = "paired"
  )
  ceiling(out$n)     # participants needed (pairs)
})

## ---- H2: treatment > control (two-sample) ----
h2_required_per_group <- sapply(ds, function(d) {
  out <- pwr.t.test(
    d = d,
    sig.level = 0.05,
    power = 0.80,
    type = "two.sample"
  )
  ceiling(out$n)   # n per group
})

h2_required_total <- 2 * h2_required_per_group

results_table <- tibble(
  Effect_Size_d = ds,
  H1_N_required_paired = h1_required,
  H2_N_required_per_group = h2_required_per_group,
  H2_N_required_total = h2_required_total
)

print(results_table)
```

# Minimally Detectable Effect Sizes

Given the actual sample ($N = 54$ total; $\approx 27$ per group), we compute the MDES at 80% power.

```{r mdes}
## ---- Minimally Detectable Effects ----

# H1 MDES
h1_mdes <- pwr.t.test(
  n = 54,
  sig.level = 0.05,
  power = 0.80,
  type = "paired"
)$d

# H2 MDES
h2_mdes <- pwr.t.test(
  n = 27,
  sig.level = 0.05,
  power = 0.80,
  type = "two.sample"
)$d

mdes_table <- tibble(
  Contrast = c("H1 (paired)", "H2 (two-sample)"),
  N = c(54, 27),
  MDES_d = c(h1_mdes, h2_mdes)
)

print(mdes_table)
```

The study has sufficient power to detect medium-to-large effects ($d \geq 0.5$) for H1, but is underpowered for small effects. For between-group comparisons (H2), the MDES is notably larger given the smaller per-group sample.
